### 2023-02-13 article.tex
Review

Additionally, there is an interdependency of scholarly findings here:
the SNP has received its name based on a computational experiment.
The experiment that concluded the usefulness of that SNP, 
as well as this experiment, are based on DNA sequences.
DNA sequences, however, are (nowadays) not simply read, 
yet are the result of a complex computational analysis,
with its own dedicated field of research.
Both studies assumed a correctly calculated DNA sequence.
This means that if the DNA sequence analysis contained a software bug, 
this study may be invalidated. 
Additionally, the result of this study may be used in follow-up studies,
that assume the result to be correctly calculated: one paper's conclusion
is the next paper's assumption.

@@

For computational science, it may appear to be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, 
only 18\% of 180 studies are easily reproducible \cite{stodden2018empirical}.
To some, it appears that 
the academic culture to reproduce results 
has been lost over time \cite{peng2011reproducible},
with labs that embrace reproducibility (for example, \cite{barba2016hard})
being the exception.
One suggested way forward is to make the reproduction of 
research a minimal requirement for publication (\cite{peng2011reproducible}).

@@

Code holds the ground truth of an experiment; it does the actual work.
The more complex the computation pipeline is, the easier it is
to have a mismatch between the article (that describes what the
code does) and the code (that actually does the work).
It is easy (tempting?) to overlook how easy it is to have a mismatch
between paper and code.

@@

It is the paper that accompanies the code,
as it is the code that generates the results.
As humans are fallible and code gets bigger, the likelihood of
a mismatch between the English paper and the code increases.
We know that the most common errors are 
simple (and may legitimately fear that those simple errors are common) 
(\cite{baggerly2009deriving, vable2021code}),
and this example is not too far off from the
examples given in that paper.

@@

Code is rarely preserved (\cite{barnes2010publish}).
If it is, keeping it running is even more rare.
This section discusses the preservation of code for a short, medium
and long term.

@@

To preserve code for a short term, 
a code hosting website is a good first step.
A code hosting website is a website where 
its users can create dedicated pages (called 'repositories')
for a project, upload code and interact with that code.
There are multiple code hosting websites, 
with GitHub being the most popular one (\cite{cosentino2017systematic}).
For example, the \LaTeX \hspace{5mm} code of this chapter is hosted on GitHub 
as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
The use of code hosting websites
has increased strongly (\cite{russell2018large}),
accommodates collaboration (\cite{perez2016ten})
and improves transparency (\cite{gorgolewski2016practical}),
due to its inherent computer-friendliness.
See \cite{cosentino2017systematic} for an extensive overview of
research conducted on GitHub.

\iffalse
An example from the author is the website (\cite{bbbqarticleissue157}),
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
as shown in Figure \ref{fig:bbbqarticleissue157}.
\fi

@@

Hosted code commonly keeps a history of file changes,
a so-called commit history,
where a commit is the description of the change, 
as well as the changes in one or more files.
For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
It is a general recommendation to put version control
on all human-produced data (\cite{wilson2014best}),
as well as openly working on the code from the start (\cite{jimenez2017four}).
Half of 56 Science magazine corresponding authors 
have such a version control mechanism (\cite{stodden2018empirical}).

% \paragraph{Metrics help assess the honest division of labour}

Most code hosts, among others, GitHub and GitLab,
automatically keep track of certain metrics.
One such metric is the amount of code each contributor has made.
In that way, authors involved in the writing of the software
are honestly acknowledged. 
\iffalse
Figure \ref{fig:daisie_contributors}
shows the number of commits the authors of DAISIE (\cite{etienne2020daisie})
have made to its code.
\fi
The distribution of contributed code per author 
is useful metadata, that can automatically
be acquired when using code hosting website.

@@

A code host, however, will not actively preserve the uploaded code:
if the code breaks tomorrow, it will go unnoticed.

@@

To preserve code for a longer timespan, 
continuous integration ('CI') allows one to 
verify if code still works and, if not, to be notified on the need for an update.
Things change: from operating systems, to libraries that code uses,
to the code itself. 
Embracing those changes is recommended (\cite{beck2000extreme})
and useful code does so.

@@

Some code hosts, among others, GitHub, 
allow the researcher/user to start software scripts upon uploading a change.
This is useful to determine the reproducibility of the code: 
one such script should run the code (on trivial data) 
to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediately see which small change was the cause of this.
Using a CI service (such as GitHub Actions for GitHub), 
is known to significantly 
increase the number of bugs exposed (\cite{vasilescu2015}) and increases
the speed at which new features are added (\cite{vasilescu2015}).
Using CI opens up the possibility to informally annotate features of
a repository. 
\iffalse
For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a \LaTeX document (i.e. this paper)
could be built.
\fi

@@

@@

However, for code to work longer, one can put the code and its dependencies in
a container, that should be best in standing the test of time,
as the degradation of software is a known feature for nearly 
four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
here a more modern (and in this context superior)
definition is used: software collapse (\cite{hinsen2019dealing}),
in which software fails due to dependencies on other 
software.
Where that paper recommends using older software foundations 
with multiple implementations (\cite{hinsen2019dealing}),
this paper recommends a modern solution that potentially has more 
longevity:
to publish the code installed in a virtual environment
(including the dependencies used at that time), as described below.

Such tools combine great with CI services.

Using CI does not prevent code from breaking. Instead,
it allows one to be the first to know if code breaks.

To preserve code for the longest time, 
both code and its dependencies are put into a so-called container.

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
(however, that article was written before virtualisation existed).
Virtualisation allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity.
Due to access rights, Singularity is the only viable option to run
an experiment on a 
high performance computing (HPC) environment (e.g. a computer cluster),
as are commonly needed in the field of genetic epidemiology experiments.

To repeat features of useful paradata: paradata should be comprehensive, 
documented appropriately, documented in co-evolution with the data,
and the data should be friendly to computers (\cite{huvila2022improving}).
In this section, these ideal properties of paradata are applied to code.
However, code has multiple uses, which have different requirements
for being useful paradata.

Code can be used to reproduce, to replicate and to 
extend a computational experiment (\cite{benureau2018re}) and
each of these uses has different requirements for the code to be useful paradata.
In this section, the recommendations how to make paradata useful
is applied to code and compared to existing best practices in software 
development, for the intended use of the code.

For code to be ideal paradata, it must be usefully 
documented (\cite{huvila2022improving}).
What is 'useful' is determined by its intended use: 
to reproduce, replicate or extend (see section \ref{sec:making-code-useful-paradata}).
The amount of documentation is, however, also determined
by the recommended practices for writing code in general.

For purposes of reproducing (same code, same data) code,
it should at least be documented how the code is run.
However, 57\% out of 56 Science papers with obtainable code (in 
total there were 180 papers) have documented their code well enough to run it 
and to be understood what it ought to do (\cite{stodden2018empirical}).

However, when an independent researcher can redo an experiment, 
it is possible (even ideal!) that no code is read at all.
Within that context, it could argued that no further documentation is needed.
However, all code in general should be documented 
'adequately' (\cite{peng2006reproducible}),
ideally writing code in such a way that it becomes 
self-explanatory (\cite{wilson2014best})
and for the remaining code to document the reasons behind it, 
its design and its purpose (\cite{wilson2014best}).

For code to be ideal paradata, 
its creation and documentation need to align, 
\emph{not least because they shape each other} (\cite{huvila2022improving}).
This emphasised part of the quote resonates strongly with the idea of 
literate programming (\cite{knuth1984literate}),
in which documentation and code are developed hand-in-hand.
Literate programming is the practice of writing a document mixed with code that 
will be run.
Contemporary examples of this idea are, among others, 
vignettes (\cite{wickham2015r}) for the R programming language (\cite{r}) 
and Jupyter notebooks (\cite{wang2020assessing})
for the Julia (\cite{Julia-2017}), Python (\cite{van1995python}) and R (\cite{r})
programming languages.
rOpenSci, a community that, among others, reviews programming
code (\cite{ram2013ropensci,ram2018community}), is 
one example of where literate programming is enforced:
code must have extensive runnable examples
as part of the documentation (\cite{ropensci_2021_6619350}).
Developing code and documentation at the same time 
leads to both better code and documentation (\cite{reenskaug1989environment}).
In general, when developing software, 
it is recommended to 
to write documentation while writing software,
as well as to include many examples (\cite{lee2018ten}).

Code should be distributed in standard ways (\cite{peng2006reproducible})
(see section \ref{sec:code-availability}
for the general availability of code).
The standard way to publish code is to use a code hosting website
(see subsection \ref{subsec:code-hosting}).

Code must be more extensive when it is (intended to be) 
used on different data,
as then 'code must act as a teacher for future developers' (\cite{sadowski2018modern}).
Error handling is one of the mechanisms to do so.
One example is when doing a t-test:
to compare the means of two distributions, each distribution
must have at least two values
and, if not, the programming language should signal the error clearly.
In genetic epidemiology, it is common to have incomplete or missing data,
so analyses should take into account that there are not always enough
observations to do a t-test.
The analytical code should clearly signal this problem.

Coding errors are extremely common 
and contribute to the reproducibility crisis in science (\cite{vable2021code}).
Testing, in general, is an important mechanism to ensure
the correctness of code (an interesting example is \cite{rahman2020exploratory}
showing bugs in scientific software on the COVID-19 pandemic).

The percentage of (lines of) code tested is called the code coverage.
Code coverage correlates with code quality (\cite{horgan1994,del1995correlation}). 
The non-profit organisation rOpenSci (\cite{ram2018community}),
which peer-reviews code,
has made it a prerequisite to have a code coverage of 100\%.
CI opens up the possibility to informally annotate the
code coverage. 
For example, build badges
\iffalse
(e.g. as shown in Figure \ref{fig:badge_codecov})
\fi
are visual indications on a website that can, among others,
show a certain R package's code has a 100\% code coverage.

% \paragraph{Test-driven development}

Having a 100\% code coverage is not uncommon,
as there is a common software development practice that 
strives for this, called test-driven development ('TDD'). 
TDD is a software development strategy in which tests are written
before new 'real' code is written.
TDD improves code quality (\cite{alkaoud2018quality,janzen2006test})
and writing comment and documentation as part of the process
is easy (\cite{shmerlin2015document}).

It is considered good practice
to add a software license (\cite{jimenez2017four}),
so that it is clear that the software can be reused.
About two-thirds of 56 Science magazine corresponding authors 
that shared their code supply a software license (\cite{stodden2018empirical}).

More than half of 315 scientists 
have their code rarely or never reviewed (\cite{vable2021code}),
although code reviews are known to
accelerate learning of the developers,
improve the quality of the code
and resulting to an experiment that is likelier to be reproducible
(\cite{vable2021code}).
Within software development best-practices, 
code reviews are recommended as well (\cite{wilson2014best}) 
The observation that code is reviewed, makes it likelier to be correct.

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.

% \paragraph{Runnable code allows others to re-use an experiment}

Containers allow a computation experiment to be highly re-usable,
as any scientist can work with it and tweak code.

For those 22 (out of 180) Science magazine corresponding authors 
that supplied code and that code was estimated to be 
reproducible, one in three had the results reproduced with 
minimal or no effort (\cite{stodden2018empirical}).

@@

For research to be reproducible, one ideally has access to
both the data used and the code.
In some fields, such as genetic epidemiology, the data are
sensitive, hence cannot be released,
yet there are methods being devised to run code on sensitive
data with assured privacy (\cite{zhang2016review,azencott2018machine}).

Next to the code, it is the data used in an experiment 
that must be made available for an experiment 
to be called 'reproducible' (\cite{peng2006reproducible}).
When the input data are sensitive, however, and thus cannot be shared with
an independent researcher, one thus cannot reproduce the
experiment.

The solution for this problem is simple:
supply a simulated (also called 'analytical' (\cite{peng2006reproducible})) dataset
together with the code.
In the case of genetic epidemiology this would mean
simulated genotypes and associated phenotypes,
as can be done with the \verb|plinkr| R package (\cite{plinkr}).
This solution, however, is known to all developers that follow
good practices when writing code, as public (i.e. simulated) 
datasets are needed for automatic CI testing of code.
One extra benefit of simulated data is that these can be used
as a benchmark, as slightly different analyses should give 
similar conclusions.

@@

One needs to be trained to know the best practices,
before these can be applied. 
Articles that suggest these best practices, however, claim that this initial
investment pays off.
Code reviews are a good way to accelerate the 
learning of team members (\cite{vable2021code}).

@@

Code needs maintenance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. 
Code that will stand the test of time perfectly 
is deemed 'impossible' (\cite{benureau2018re}).

@@

There is another problem that authors of academic code face,
which is that they may be contacted with a user 
asking for technical report 
(against which a 'no' is just fine (\cite{barnes2010publish}))
or -more relevant to the research- sending in a bug report.
One solution for the author is to ignore such emails (as is done in a third
of all cases (\cite{teunis2015corresponding})):
it can be argued that no energy should be wasted on published code
and work on something new instead 
(see \cite{barnes2010publish} for a better way to deal with the problem).

@@

However, upon the find of a new bug, the question is:
does all the research that use that software still 
result in the same conclusions?
To prevent authors choosing to hide bug reports, 
a code hosting website usually has a public place to report bugs,
to increase transparency in the handling of bug reports.
Should there be a retraction notice on code?

@@

There are some problems with providing a container with a runnable container.
First, a container is several gigabytes and need to be stored somewhere
online. 
Ideally, for Singularity, this would be at the main container hosting
site called Syslabs. Using Syslabs is free upon registration, 
yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
be entered, such as the URL of the article, nor the URL of hosted code.

% \paragraph{Code in container}

Here are multiple suggestions to make containers with runnable code FAIR.
To make such containers findable, the most common 
environments (Syslabs and DockerHub) to upload containers must be indexed.
Uploading a container to those environments is, however, limited by
a restricted amount of free storage, with the possible consequence that
those containers are uploaded somewhere where they escape proper indexing.
Having containers uploaded at the existing websites ensures they are
accessible, interoperable and reusable.
One could argue however, that the scientific community need its own,
similar website to upload containers, as the metadata for these
sites are limited to free-text comments and tags (DockerHub)
or not metadata at all (Syslabs).

% \paragraph{The drawbacks of publishing code}

Publishing code may be disadvantageous for an author.
For science, code should be published, 
this allows reproducible research 
(see \cite{haibe2020importance} for a tragic example).
For an author, publishing code alongside an experiment opens up
the possibility to receive questions regarding that code.
Note, however, that not publishing code may put 
oneself in the focus of attention
and -after much effect by others reproducing an incorrect result-
at the cost of a scientific career (\cite{baggerly2009deriving}).

@@

For the preservation of code, these are the recommendations:

>>>>

Additionally, there is an interdependency of scholarly findings here:
the SNP has received its name based on a computational experiment.
This earlier experiment that concluded the usefulness of that SNP, 
is based on some DNA sequences.
This experiment is based on assumed DNA sequences.
DNA sequences, however, are (nowadays) not simply read, 
yet are the result of a complex computational analysis instead,
with its own dedicated field of research.
Both studies assumed a correctly calculated DNA sequence.
This means that if the DNA sequence analysis contained a software bug, 
this study may be invalidated. 
Additionally, the result of this study may be used in follow-up studies,
that assume the result to be correctly calculated: one paper's conclusion
is the next paper's assumption.

@@

For computational science, it may appear to be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, 
only 18\% of 180 computational studies 
are easily reproducible (\cite{stodden2018empirical}).
To some, it appears that 
the academic culture to reproduce results 
has been lost over time \cite{peng2011reproducible},
with labs that embrace reproducibility (for example, \cite{barba2016hard})
being the exception.
One suggested way forward is to make the reproduction of 
research a minimal requirement for publication (\cite{peng2011reproducible}).

@@

@@

Code holds the ground truth of an experiment; it does the actual work.
The more complex the computation pipeline is, the easier it is
to have a mismatch between the article (that describes what the
code does) and the code (that actually does the work).
The moment these two disagree, it is the code that is true.

@@

Code is rarely preserved (\cite{barnes2010publish}).
This section discusses the preservation of code for a short, medium
and long term.

@@

To preserve code for a short term, 
a code hosting website is a good first step.
A code hosting website is a website where 
its users can create dedicated pages (called 'repositories')
for a project, upload code and interact with that code.
There are multiple code hosting websites, 
with GitHub being the most popular one (\cite{cosentino2017systematic}).
The use of code hosting websites
has increased strongly (\cite{russell2018large}),
accommodates collaboration (\cite{perez2016ten})
and improves transparency (\cite{gorgolewski2016practical}),
due to its inherent computer-friendliness.
See \cite{cosentino2017systematic} for an extensive overview of
research conducted on GitHub.

\iffalse
An example from the author is the website (\cite{bbbqarticleissue157}),
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
as shown in Figure \ref{fig:bbbqarticleissue157}.
Also, the \LaTeX \hspace{5mm} code of this chapter is hosted on GitHub 
as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
\fi

@@

Hosted code commonly keeps a history of file changes,
This means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
The version control system keeps track of who-did-what transparantly.
\iffalse
Figure \ref{fig:daisie_contributors}
shows the number of commits the authors of DAISIE (\cite{etienne2020daisie})
have made to its code.
\fi
It is a general recommendation to put version control
on all human-produced data (\cite{wilson2014best}),
as well as openly working on the code from the start (\cite{jimenez2017four}).
Half of the published code has such a version control 
system (\cite{stodden2018empirical}).

@@

The degradation of software is a known feature for nearly 
four decades.
This is called 'bit rot' by \cite{steele1983hacker},
or 'software collapse' by \cite{hinsen2019dealing},
in which software fails due to dependencies on other 
software.
Using a code hosting website, 
that only passively stores code,
ignores this problem.

@@

To preserve code for a longer timespan, 
it needs to be embraced that software degrades (\cite{beck2000extreme}).
Continuous integration ('CI') allows one to 
verify if code still works and, if not, to be notified.

@@

Some code hosts,
allow a user to trigger specialised code upon uploading a change,
called a CI script.
Such a CI script typically builds and tests the code.
This practice is known to significantly 
increase the number of bugs exposed (\cite{vasilescu2015}) and increases
the speed at which new features are added (\cite{vasilescu2015}).
CI can be scheduled to run on a regular basis and
notify the user directly when the code has broken down.

To preserve code for the longest time, 
both code and its dependencies can be put into a so-called container.

The most reproducible way of submitting the code of an experiment,
is to put all code with all the (software) dependencies
in a file that acts as a virtual computational environment,
called a 'virtual container' (from now on: 'container').
Such a container is close to the golden standard of reproducible research
as suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
.

Useful paradata, in general, are (1) comprehensive,
(2) documented appropriately, (3) documented in co-evolution with the data,
and (4) friendly to computers (\cite{huvila2022improving}).
In this section, these ideal properties of paradata are applied to code.

However, code has multiple uses,
as code can be used to (1) reproduce, (2) replicate, or (3)
extend a computational experiment (\cite{benureau2018re}).
Depending on the intended use of the code,
there are different requirements for code being useful paradata.

For code to be ideal paradata, it must be usefully 
documented (\cite{huvila2022improving}).

For purposes of reproducing code,
it should at least be documented how to run the code and what it ought to do.
Although this may be obvious, only 57\% out of 56 Science papers 
with obtainable code (in total there were 180 papers) 
do so (\cite{stodden2018empirical}).

When it can be found out how how an experiment is run, 
it is possible (even ideal!) that no code is read at all.
Within that context, it could argued that no (further) documentation is needed.
However, all code in general should be documented 
'adequately' (\cite{peng2006reproducible}),
ideally writing code in such a way that it becomes 
self-explanatory (\cite{wilson2014best})
and for the remaining code to document the reasons behind it, 
its design and its purpose (\cite{wilson2014best}).

For code to be ideal paradata, 
its creation and documentation need to align, 
\emph{not least because they shape each other} (\cite{huvila2022improving}).
This emphasised part of the quote resonates strongly with the idea of 
literate programming (\cite{knuth1984literate}),
in which documentation and code are developed hand-in-hand.
Literate programming is the practice of writing code and 
documentation in the same file.
Contemporary examples of this idea are, among others, 
vignettes (\cite{wickham2015r}) for the R programming language (\cite{r}) 
and Jupyter notebooks (\cite{wang2020assessing})
for the Julia (\cite{Julia-2017}), Python (\cite{van1995python}) and R (\cite{r})
programming languages.
rOpenSci, a community that, among others, reviews programming
code (\cite{ram2013ropensci,ram2018community}), is 
one example of where extensive documentation is mandatory
and all code must have examples (that are actually run)
as part of the documentation (\cite{ropensci_2021_6619350}).
In general, when developing software, 
it is recommended to 
to write documentation while writing software,
as well as to include many examples (\cite{lee2018ten}),
as this leads to both better code and documentation (\cite{reenskaug1989environment}).

% \paragraph{Code should be distributed in standard ways}

Code should be distributed in standard ways (\cite{peng2006reproducible}),
as is done by using a code hosting website
(see subsection \ref{subsec:code-hosting}).

Code must be more extensive when it is (intended to be) 
used on different data,
as then 'code must act as a teacher for future developers' (\cite{sadowski2018modern}).
Error handling is one of the mechanisms to do so.
In genetic epidemiology, it is common to have incomplete or missing data,
so analyses should take this into account with clear error messages.

Coding errors are extremely common (\cite{baggerly2009deriving, vable2021code})
and contribute to the reproducibility crisis in science (\cite{vable2021code}).
Testing, in general, is an important mechanism to ensure
the correctness of code.
One clear example is \cite{rahman2020exploratory}, 
showing bugs in scientific software on the COVID-19 pandemic.

% \paragraph{TDD}

Testing is so important that it is at the
heart of a software development methodology called 'Test-Driven 
Development' ('TDD'), in which tests are written
before the 'real' code.
TDD improves code quality (\cite{alkaoud2018quality,janzen2006test})
and it is easy to integrate 
the writing of documentation as part of the TDD cycle (\cite{shmerlin2015document}).

The percentage of (lines of) code tested is called the code coverage.
Code coverage correlates with code quality (\cite{horgan1994,del1995correlation}).
and, due to this, having a code coverage of (around) 100\% 
is mandatory to pass a code peer-review
by rOpenSci (\cite{ram2018community}).
When CI is activated, 
the code coverage of a project can be shown on the repository's website
\iffalse
(e.g. as shown in Figure \ref{fig:badge_codecov})
\fi
.

\iffalse
CI opens up the possibility to informally annotate features of
a repository. 
For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a \LaTeX document (i.e. this paper)
could be built.
\fi

@@

It is considered good practice
to add a software license (\cite{jimenez2017four}),
so that it is clear that the software can be reused.
Although this may seem trivial, 
only two-thirds of 56 computational experiments 
supply a software license (\cite{stodden2018empirical}).

Code reviews are recommended by software development 
best-practices (\cite{wilson2014best}).
However, more than half of 315 scientists 
have their code rarely or never reviewed (\cite{vable2021code}),
although code reviews are known to
accelerate learning of the developers,
improve the quality of the code
and resulting to an experiment that is likelier to be reproducible
(\cite{vable2021code}).

The most reproducible way of submitting the code of an experiment,
is by providing the code with all its (software) dependencies 
in a container.

@@

@@

Next to the code, it is the data used in an experiment 
that must be made available for an experiment 
to be called 'reproducible' (\cite{peng2006reproducible}).
In some fields, such as genetic epidemiology, the data are
sensitive, hence cannot be released, thus one cannot reproduce 
an experiment.
To solve this problem in the future, there are some 
interesting methods being developed to run code on sensitive
data with assured privacy (\cite{zhang2016review,azencott2018machine}).

To alleviate the problem today,
a developer should supply a simulated 
(also called 'analytical' (\cite{peng2006reproducible})) dataset
together with the code.
This simulated data is needed to run tests, as is 
part of the TDD methodology.
In the case of genetic epidemiology this would mean
simulated genotypes and associated phenotypes,
as can be done with the \verb|plinkr| R package (\cite{plinkr}).
One extra benefit of simulated data is that these can be used
as a benchmark, as slightly different analyses should give 
similar conclusions.

@@

To know these best practices, one needs to be trained. 
Articles that suggest these best practices (such as this one), 
claim that this initial investment pays off.
Code reviews are a good way to accelerate the 
learning of team members (\cite{vable2021code}).

@@

Code needs maintenance,
as code that will stand the test of time perfectly 
is deemed 'impossible' (\cite{benureau2018re}).
CI can help a maintainer to be notified when the code breaks,
where the use of containers may slow down time,
as an entire computational environment is preserved.

% \paragraph{Uploading code may feel like a risk}

Uploading code, preferably to a code hosting website, 
may feel like a risk, as all code can be seen and scrutinised.
However, not publishing code may put 
oneself in the focus of attention
and -after much effect by others reproducing an incorrect result-
at the cost of a scientific career (\cite{baggerly2009deriving}).

@@

When the author of code can be contacted,
there will be users asking for technical support 
(against which a 'no' is just fine (\cite{barnes2010publish}))
One solution for the author is to ignore such emails (as is done in a third
of all cases (\cite{teunis2015corresponding})):
it can be argued that no energy should be wasted on published code
and work on something new instead.
However, see \cite{barnes2010publish for a better way to deal with this problem.

@@

When the author of code can be contacted,
users will send in bug reports.
If the bug is severe enough, the question arises
if all the research that use that code still 
results in the same conclusions.
One such bug is described in \cite{eklund2016cluster},
with 40,000 studies using that incorrect code.
These reports could be ignored to work on something new.

@@

One problem that containers have, is their size:
they can be several gigabytes big. 
This makes the distribution of these containers harder.
Ideally, containers are stored online and distributed in
standardised ways.
Although progress is being made, 
there is no way to do so for all container types.
Additionally, probably due to their novely,
container hosting sites lack metadata.

@@

For the preservation of code, these recommendations are made:
